---
title: "03_ComparePopulations"
author: "Maria Jose Herrera"
date: "6/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library("tidyverse")
library("ggplot2")
library("reshape2")
```


Define mortgage prisoners: 1) self-employed 
                           2) interest-only mortgages
                           3) spending >33% income on mortgage repayment 
                              OR report having had a hard time making mortgage repayments over the last 12 months

# Load data
```{r}
# List of files to open (final mtg prisoner + conditions individually for desc stats)
files <- list(
  # all mtg prisoners
  "/Users/mariajoseherrera/Documents/LSE_new/03_Dissertation/Understanding_Society/df_all_incl_mtgprisoners.rds",
  # those that meet condition 1: self-employed
  "/Users/mariajoseherrera/Documents/LSE_new/03_Dissertation/Understanding_Society/df_cond1.rds",
  # those that meet condition 2: interest-only mtg
  "/Users/mariajoseherrera/Documents/LSE_new/03_Dissertation/Understanding_Society/df_cond2.rds",
  # those that meet condition 3_1: spending > 33% of monthly income on mtg repayment
  "/Users/mariajoseherrera/Documents/LSE_new/03_Dissertation/Understanding_Society/cond3_1.rds",
  # those that meet condition 3_2: report having a hard-time making mtg repayments over the last 12 months
  "/Users/mariajoseherrera/Documents/LSE_new/03_Dissertation/Understanding_Society/cond3_2.rds",
  "/Users/mariajoseherrera/Documents/LSE_new/03_Dissertation/Understanding_Society/df_mtgprisoners.rds") 

# Read in data
data_files <- lapply(files, readRDS)
# Assign names to each (based on order above)
df_all <- data_files[[1]]
df_cond1 <- data_files[[2]]
df_cond2 <- data_files[[3]]

df_cond3_1 <- data_files[[4]]
df_cond3_2 <- data_files[[5]]
df_mtgprisoners <- data_files[[6]]


```

# Categorize variables
```{r}
# Numeric
 numeric_vars <- c("fihhmnnet1_dv", # total monthly income, no deductions
                   "hscost", # original purchase price of property
                   "hsval", # value of property: home owners
                   "mglife", # years left to pay: mortgage
                   "hhsize", # household size
                   "mgold", # initial mortgage	
                   "houscost1_dv", # monthly housing cost /in/cluding mortgage principal payments 
                   "houscost2_dv", # monthly housing cost /ex/cluding mortgage principal payments
                   "fimnnet_dv", # Own total estimated net monthly income
                   "fimnlabnet_dv", # Own total estimated net monthly income from labour
                   "scend_dv", # school leaving age (xwavedat)
                   "sf12mcs_dv", # SF-12 mental health component score, derived
                   "sf12pcs_dv", # Sf-12: physical health component score, derived
                   "scghq1_dv", # Subjective wellbeing: Likert
                   "scghq2_dv", # Subjective wellbeing: Caseness
                   "age_dv", # Age at time of interview, derived
                   "nchild_dv", # Number of kids in the household
                   "prfitba", # Total personal income annually
                   "wavenumber") 



# Categorical
cat_vars <- c("hsyr04", # year mortgage began
              "mgtype", # mortgage type
              "hiqual_dv", #Highest qualification
              "jbsoc00_cc", # Standard socio-economic classification (SOC 2000) of current job; condensed 3 digit version
              "jbnssec8_dv", # Current job (eight class ns-sec)
              "jbnssec5_dv", # Current job (five class ns-sec)
              "jbnssec3_dv", # Current job (three class ns-sec)
              "health", # Long-standing illness or disability
              "country", # Country in the UK
              "gor_dv", # Region in the UK
              "urban_dv", # Urban or rural area, derived
              "sex_dv", # sex, derived
              "marstat", # Marital status
              "jbstat", # Employment status
              "jbsemp", # Employed or self-employed
              "racel_dv", # Ethnic group (self-reported) (indresp)
              "ethn_dv", # Ethnic group - derived from multiple sources (indresp)
  # Need to re-run code to get this var #   "j1soc00_cc", # Standard Socio-economic Classification (SOC 2000) of first job after leaving full-time education. Condensed three-digit version (xwavedat)
              "maid", # mother's ethnic group (xwavedat)
              "macob", # mother's country of birth (xwavedat)
              "maedqf", # mother's educational qualification when respondent was aged 14 (xwavedat)
              "masoc00_cc", # Standard Occupational Classification 2000 of mother's job when respondent was aged 14 (xwavedat)
              "paid", # father's ethnic group (xwavedat)
              "pacob", # father's country of birth (xwavedat)
              "paedqf", # father's educational qualification when respondent was aged 14 (xwavedat)
              "pasoc00_cc", # Standard Occupational Classification 2000 of father's job when respondent was aged 14 (xwavedat)
              "wavenumber") 


```


# Calculate number of people per household (NB THERES A VARIABLE THAT DOES THIS SO THIS IS USELESS)
## Applying all conditions
```{r}
# For entire population
count_perhh <- df_all %>% count(hidp)

all <- summary(count_perhh$n) # mean is 1.862, median is 2, min is 1, max is 11

# For non-mortgage prisoners
nonmtg_count_perhh <- df_all[df_all$is_mtgprisoner == 0, ] %>% count(hidp)
non_mtg <- summary(nonmtg_count_perhh$n) # mean is 1.769, median is 2, min is 1, max is 8

# For mortgage prisoners
mtg_count_perhh <- df_all[df_all$is_mtgprisoner == 1, ] %>% count(hidp)
all_mtg_prisoners <- summary(mtg_count_perhh$n) # mean is 1.544, median is 1, min is 1, max is 9

## TODO: Do these descriptive stats by condition (why is mean lower than non-mortgage prisoners?

```

## Apply condition 1
```{r}
# For condition 1: self-employed
cond1_count_perhh <- df_cond1 %>% count(hidp)
cond1 <- summary(cond1_count_perhh$n) # mean number of people in a hh is 1.104, median is 1, min is 1, max is 4

ggplot(count_perhh) +
  geom_bar(aes(x = n))
```

## Apply condition 2
```{r}
# For condition 2: interest-only mortgages
cond2_count_perhh <- df_cond2 %>% count(hidp)
cond2 <- summary(cond2_count_perhh$n) # mean number of people in a hh is 1.837, median is 2, min is 1, max is 11

ggplot(count_perhh) +
  geom_bar(aes(x = n))
                           
```

## Apply condition 3_1
```{r}
# For condition 3_1: spending >33% income on mortgage repayment 
cond31_count_perhh <- df_cond3_1 %>% count(hidp)
cond3_1 <- summary(cond31_count_perhh$n) # mean number of people in a hh is 1.574, median is 1, min is 1, max is 5                 

ggplot(count_perhh) +
  geom_bar(aes(x = n))
```


## Apply condition 3_2
```{r}
# For condition 3_2: report having had a hard time making mortgage repayments over
cond32_count_perhh <- df_cond3_2 %>% count(hidp)
cond3_2 <- summary(cond32_count_perhh$n) # mean number of people in a hh is 1.889, median is 2, min is 1, max is 8

ggplot(count_perhh) +
  geom_bar(aes(x = n))

```


# Compare household size

```{r}
df_summary <- as.data.frame(do.call(cbind, lapply(list("All" = count_perhh$n, "All non-mtg prisoners" = nonmtg_count_perhh$n, "All mtg. prisoners" = mtg_count_perhh$n,"Cond. 1" = cond1_count_perhh$n, "Cond. 2" = cond2_count_perhh$n, "Cond. 3_1" = cond31_count_perhh$n, "Cond. 3_2" = cond32_count_perhh$n), summary)))

df_summary <- as.data.frame(t(df_summary))
df_summary

saveRDS(df_summary, "/Users/mariajoseherrera/Documents/LSE_new/03_Dissertation/Understanding_Society/hhsize_summary.rds")

```


# Plot all vars
## Entire population
```{r}
melt_df_all <- melt(df_all)

ggplot(data = melt_df_all, aes(x = value)) + 
stat_density() + 
facet_wrap(~variable, scales = "free")

```

## Just mtg prisoners
### All vars
```{r}
melt_df_mtgprisoners <- melt(df_mtgprisoners) #df of unique mtg prisoners

ggplot(data = melt_df_mtgprisoners, aes(x = value)) + 
stat_density() + 
facet_wrap(~variable, scales = "free")

```

### Categorical variables - all time
```{r}
df_categorical <- select(df_mtgprisoners, one_of(cat_vars))

# Get frequency
cat_freq <- mapply(table, df_categorical)

# Get percentages
lapply(cat_freq, prop.table)

# Summary tables
mapply(summary, df_categorical)

# Visualizations
melt_df_categorical <- melt(df_categorical)
ggplot(melt_df_categorical, aes(x = value)) +
  geom_bar() +
  facet_wrap(~variable, scales = "free") +
  theme_bw()
  

```

### Categorical variables - after legis
```{r}
# Load df
df_post <- readRDS("/Users/mariajoseherrera/Documents/LSE_new/03_Dissertation/Understanding_Society/df_mtgprisoners_postlegis.rds")

```

```{r}
df_categorical_post <- select(df_post, one_of(cat_vars))

# Get frequency
post_cat_freq <- mapply(table, df_categorical)

# Get percentages
lapply(post_cat_freq, prop.table)

# Summary tables
mapply(summary, df_categorical_post)

# Visualizations
melt_df_categorical_post <- melt(df_categorical_post)
ggplot(melt_df_categorical_post, aes(x = value)) +
  geom_bar() +
  facet_wrap(~variable, scales = "free") +
  theme_bw()

```

### Numeric variables - all time
```{r}
df_numeric <- select(df_mtgprisoners, one_of(numeric_vars))

# TODO: run wave data collection again to collect "hhsize" information (NEED THIS BC THERE MIGHT BE KIDS IN A HH)

# Get frequency
num_freq <- mapply(table, df_numeric)

# Get percentages
lapply(num_freq, prop.table)

# Summary tables
mapply(summary, df_numeric)

# Visualizations
melt_df_numeric <- melt(df_numeric)
ggplot(melt_df_numeric, aes(x = value)) +
  stat_density() +
  facet_wrap(~variable, scales = "free") +
  theme_bw()
```

### Numeric variables - post legis
```{r}


```


