---
title: "04_HierarchicalClustering"
author: "Maria Jose Herrera"
date: "7/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("tidyverse")  # data manipulation
library("cluster")    # clustering algorithms
library("factoextra") # clustering visualization
library("dendextend") # for comparing two dendrograms

# source: https://uc-r.github.io/hc_clustering
# source for processing/choosing number of clusters: https://towardsdatascience.com/hierarchical-clustering-on-categorical-data-in-r-a27e578f2995
```

# Load data
```{r}
df_postlegis <- readRDS("/Users/mariajoseherrera/Documents/LSE_new/03_Dissertation/Understanding_Society/df_mtgprisoners_postlegis.rds")

df_postlegis$is_mtgprisoner <- NULL

df_prelegis <- readRDS("/Users/mariajoseherrera/Documents/LSE_new/03_Dissertation/Understanding_Society/df_mtgprisoners_prelegis.rds")


```

# Prepare data

1) Rows are observations (individuals) and columns are variables
2) Any missing value in the data must be removed or estimated.
3) The data must be standardized (i.e., scaled) to make variables comparable. Recall that, standardization consists of transforming the variables such that they have mean zero and standard deviation one.[^scale]
```{r}


```

## build dissimilarity matrix using gower for cateogrical variables
```{r}
# Categorical
cat_vars <- c("hsyr04", # year mortgage began
              "mgtype", # mortgage type
              "hiqual_dv", #Highest qualification
              "jbsoc00_cc", # Standard socio-economic classification (SOC 2000) of current job; condensed 3 digit version
              "jbnssec8_dv", # Current job (eight class ns-sec)
              "jbnssec5_dv", # Current job (five class ns-sec)
              "jbnssec3_dv", # Current job (three class ns-sec)
              "health", # Long-standing illness or disability
              "country", # Country in the UK
              "gor_dv", # Region in the UK
              "urban_dv", # Urban or rural area, derived
              "sex_dv", # sex, derived
              "marstat", # Marital status
              "jbstat", # Employment status
              "jbsemp", # Employed or self-employed
              "racel_dv", # Ethnic group (self-reported) (indresp)
              "ethn_dv", # Ethnic group - derived from multiple sources (indresp)
  # Need to re-run code to get this var #   "j1soc00_cc", # Standard Socio-economic Classification (SOC 2000) of first job after leaving full-time education. Condensed three-digit version (xwavedat)
              "maid", # mother's ethnic group (xwavedat)
              "macob", # mother's country of birth (xwavedat)
              "maedqf", # mother's educational qualification when respondent was aged 14 (xwavedat)
              "masoc00_cc", # Standard Occupational Classification 2000 of mother's job when respondent was aged 14 (xwavedat)
              "paid", # father's ethnic group (xwavedat)
              "pacob", # father's country of birth (xwavedat)
              "paedqf", # father's educational qualification when respondent was aged 14 (xwavedat)
              "pasoc00_cc", # Standard Occupational Classification 2000 of father's job when respondent was aged 14 (xwavedat)
              "wavenumber") 

df_postlegis[cat_vars] <- lapply(df_postlegis[cat_vars], factor)

gower_dist <- daisy(df_postlegis, metric = c("gower"))
```



# Agglomerative clustering (AGNES) - bottom-up

(According to Berkeley source above, good at identifying small clusters)

## hclust()

### No preproccessing
```{r}
# Dissimilarity matrix
d <- dist(df_postlegis, method = "euclidean")

# Hierarchical clustering using Complete Linkage
hc1 <- hclust(d, method = "complete" )

# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)


```

### With processing
```{r}
aggl_clust_c <- hclust(gower_dist, method = "complete")
plot(aggl_clust_c, main = "Agglomerative, complete linkages")

```

```{r}
# silhouette method

```


## agnes()
```{r}
# Compute with agnes
hc2 <- agnes(df_postlegis, method = "complete")

# Agglomerative coefficient
hc2$ac


```

# Divisive clustering (DIANA) - top down

(According to Berkeley source above, good at identifying large clusters)
```{r}
divisive.clust <- diana(as.matrix(gower_dist), 
                  diss = TRUE, keep.diss = TRUE)
plot(divisive.clust, main = "Divisive")

```

# Stats
```{r}
library("fpc")

cstats.table <- function(dist, tree, k) {
clust.assess <- c("cluster.number","n","within.cluster.ss","average.within","average.between",
                  "wb.ratio","dunn2","avg.silwidth")
clust.size <- c("cluster.size")
stats.names <- c()
row.clust <- c()
output.stats <- matrix(ncol = k, nrow = length(clust.assess))
cluster.sizes <- matrix(ncol = k, nrow = k)
for(i in c(1:k)){
  row.clust[i] <- paste("Cluster-", i, " size")
}
for(i in c(2:k)){
  stats.names[i] <- paste("Test", i-1)
  
  for(j in seq_along(clust.assess)){
    output.stats[j, i] <- unlist(cluster.stats(d = dist, clustering = cutree(tree, k = i))[clust.assess])[j]
    
  }
  
  for(d in 1:k) {
    cluster.sizes[d, i] <- unlist(cluster.stats(d = dist, clustering = cutree(tree, k = i))[clust.size])[d]
    dim(cluster.sizes[d, i]) <- c(length(cluster.sizes[i]), 1)
    cluster.sizes[d, i]
    
  }
}
output.stats.df <- data.frame(output.stats)
cluster.sizes <- data.frame(cluster.sizes)
cluster.sizes[is.na(cluster.sizes)] <- 0
rows.all <- c(clust.assess, row.clust)
# rownames(output.stats.df) <- clust.assess
output <- rbind(output.stats.df, cluster.sizes)[ ,-1]
colnames(output) <- stats.names[2:k]
rownames(output) <- rows.all
is.num <- sapply(output, is.numeric)
output[is.num] <- lapply(output[is.num], round, 2)
output
}
# I am capping the maximum amout of clusters by 7
# I want to choose a reasonable number, based on which I will be able to see basic differences between customer groups as a result
stats.df.divisive <- cstats.table(gower_dist, divisive.clust, 7)
stats.df.divisive


```

# Choosing number of clusters
```{r}
# Using "Elbow" and "Silhouette" methods to identify the best number of clusters
# to better picture the trend, I will go for more than 7 clusters.
library("ggplot2")

# Elbow

# Divisive clustering
ggplot(data = data.frame(t(cstats.table(gower_dist, divisive.clust, 15))), 
  aes(x=cluster.number, y=within.cluster.ss)) + 
  geom_point()+
  geom_line()+
  ggtitle("Divisive clustering") +
  labs(x = "Num.of clusters", y = "Within clusters sum of squares (SS)") +
  theme(plot.title = element_text(hjust = 0.5))

# Agglomerative clustering,provides a more ambiguous picture
ggplot(data = data.frame(t(cstats.table(gower_dist, aggl_clust_c, 15))), 
  aes(x=cluster.number, y=within.cluster.ss)) + 
  geom_point()+
  geom_line()+
  ggtitle("Agglomerative clustering") +
  labs(x = "Num.of clusters", y = "Within clusters sum of squares (SS)") +
  theme(plot.title = element_text(hjust = 0.5))

```

# Using towards data science method for mixed data types
```{r}
# Load useful packages
library("cluster")
library("dplyr")
library("ggplot2")
library("readr")
library("Rtsne")
```

## build dissimilarity matrix using gower for cateogrical variables
```{r}
# Categorical
cat_vars <- c("hsyr04", # year mortgage began
              "mgtype", # mortgage type
              "hiqual_dv", #Highest qualification
              "jbsoc00_cc", # Standard socio-economic classification (SOC 2000) of current job; condensed 3 digit version
              "jbnssec8_dv", # Current job (eight class ns-sec)
              "jbnssec5_dv", # Current job (five class ns-sec)
              "jbnssec3_dv", # Current job (three class ns-sec)
              "health", # Long-standing illness or disability
              "country", # Country in the UK
              "gor_dv", # Region in the UK
              "urban_dv", # Urban or rural area, derived
              "sex_dv", # sex, derived
              "marstat", # Marital status
              "jbstat", # Employment status
              "jbsemp", # Employed or self-employed
              "racel_dv", # Ethnic group (self-reported) (indresp)
              "ethn_dv", # Ethnic group - derived from multiple sources (indresp)
  # Need to re-run code to get this var #   "j1soc00_cc", # Standard Socio-economic Classification (SOC 2000) of first job after leaving full-time education. Condensed three-digit version (xwavedat)
              "maid", # mother's ethnic group (xwavedat)
              "macob", # mother's country of birth (xwavedat)
              "maedqf", # mother's educational qualification when respondent was aged 14 (xwavedat)
              "masoc00_cc", # Standard Occupational Classification 2000 of mother's job when respondent was aged 14 (xwavedat)
              "paid", # father's ethnic group (xwavedat)
              "pacob", # father's country of birth (xwavedat)
              "paedqf", # father's educational qualification when respondent was aged 14 (xwavedat)
              "pasoc00_cc", # Standard Occupational Classification 2000 of father's job when respondent was aged 14 (xwavedat)
              "wavenumber") 

df_postlegis[cat_vars] <- lapply(df_postlegis[cat_vars], factor)
```


## Compute Gower distance
```{r}
# Move hidp and pidp to front of DF
df_postlegis <- df_postlegis[c(1,24,2:23,25:48)] # ONLY RUN THIS ONCE OTHERWISE OTHER IS FUCKED
```

```{r}
# Compute Gower distance
gower_dist <- daisy(df_postlegis[3:ncol(df_postlegis)], metric = c("gower")) ## looking only at mortgage prisoners after law was passed; EXCLUDE PIDP AND HIDP (columns 1 and 2)

gower_mat <- as.matrix(gower_dist)

# Print most similar respondents
df_postlegis[which(gower_mat == min(gower_mat[gower_mat != min(gower_mat)]), arr.ind = TRUE)[1, ], ]

# Print most dissimilar respondents
df_postlegis[which(gower_mat == max(gower_mat[gower_mat != max(gower_mat)]), arr.ind = TRUE)[1, ], ]
```

## Silhouette method to find optimal number of clusters
```{r}
sil_width <- c(NA)
for(i in 2:30){  # they limit it to 8 bc of "interpretability" / can look into this -- i put 20 for fun
  pam_fit <- pam(gower_dist, diss = TRUE, k = i)  
  sil_width[i] <- pam_fit$silinfo$avg.width  
}
plot(1:30, sil_width,
     xlab = "Number of clusters",
     ylab = "Silhouette Width")
lines(1:30, sil_width)

```

## Interpretation

```{r}
k <- 10 # pick number of clusters based on silhouette
pam_fit <- pam(gower_dist, diss = TRUE, k)
pam_results <- df_postlegis[3:ncol(df_postlegis)] %>%
  mutate(cluster = pam_fit$clustering) %>%
  group_by(cluster) %>%
  do(the_summary = summary(.))
pam_results$the_summary

```

## Visualization in a lower dimensional space
```{r}
tsne_obj <- Rtsne(gower_dist, is_distance = TRUE)
tsne_data <- tsne_obj$Y %>%
  data.frame() %>%
  setNames(c("X", "Y")) %>%
  mutate(cluster = factor(pam_fit$clustering))
ggplot(aes(x = X, y = Y), data = tsne_data) +
  geom_point(aes(color = cluster))

```